### 2020-02-28 analysis/paper.Rmd
replicating our own repository

---
title: "Replication in Archaeological Science: Shape as a Measure of Weapon Standardisation"
author: "Team Fresh"
date: "2/24/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

## Introduction

The traditional approach to evaluating the level of manufacturing standardization in a society has involved the comparison of standard dimensional measurements of the object through the coefficient of variation (CV) metric [1]. Computed as a simple ratio of the standard deviation to the mean, and scaled by 100 as shown below, the coefficient of variation provides an easily interpretable measure of variation that can be compared to other measurements with different mean values.

$$ CV = \frac{stdev}{mean}*100 $$

However, in the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres, a new approach is proposed. Instead of relying on dimensional comparison of different CVs, they propose using a Geomorphic Morphometric (GMM) analysis as a general measure of overall shape [2].

In this work, we are attempting to confirm that the underlying data provided is the same as that used in the original study by replicating the figure showing the pair-wise correlation of all lance dimensions across the three underlying datasets (original study Fig 8). We are also attempting to replicate the original finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis.

## Methods

The original analysis was conducted in an unspecified version of R. However, specific versions of geomorph(v3.0.6) and shapes(v1.2.3), two of the most important packages, were directly stated. Our reproduction attempt began with importing the packages explicitly identified in the paper as well as those required in the technical code.

```{r load packages, include=TRUE}
#adding this line as a git test
#install specific versions of packages called out in the paper
#require(devtools)
library(devtools)
#install_version("geomorph", version = "3.0.6", repos = "http://cran.rstudio.com")
#install_version("geomorph", version = "3.0.6", repos = "http://cran.r-project.org")
#install.packages("http://https://cran.r-project.org/src/contrib/Archive/geomorph/geomorph_3.0.6.tar.gz", repos=NULL, type="source")
library(geomorph) # geomorph 3.0.6

#install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.rstudio.com")
#install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.us.r-project.org")
library(shapes) # shapes 1.2.3

#not mentioned in the technical appendix, but required in technical code
library(shapes)

#package for lattice replication
library(GGally)

#package to support updating geomorph code for those that cannot use old package
#library(RRPP)

#Running on OSX instead of Windows advice:
# 1. Install XCode (this enables devtools)
# 2. Install XQuartz https://www.xquartz.org/
# 3. Install gfortran https://gcc.gnu.org/wiki/GFortranBinariesMacOS

```

Our next step was to set the working directory. Although a trivial step, the original code used setwd() and a static path string while our reproduction attempt opted to use the more robust here() library as shown below.

```{r set working directory, include=TRUE}
#setwd(dirname(rstudioapi::getSourceEditorContext()$path))
#getwd()
library(here)
```

Next, we read in and shaped the data obtained from the ScieceDirect archive of the paper using the same code as the original paper, slightly modified to account for our modified approach to setting the working directory.

```{r load data, include=TRUE}
# Read the appendix datafile containing a matrix of dimensional data (metric) as "lances_M" 
lances_M<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc2.csv"),header=TRUE)

# Read the datafile "lances_GMM" containing a matrix of landmark data (geometric morphometric)
#lances_GMM<-read.csv("Data/1-s2.0-S0305440318306757-mmc3.csv",header=TRUE)
lances_GMM<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc3.csv"),header=TRUE)

# Convert the 2D matrix of geometric morphometric data ("lances_GMM") into a 3D array
lances_geo<-arrayspecs(lances_GMM[,2:15],7,2) # 7 landmarks, recording 2 dimensions only (xy coordinates), ignoring the first column of lance IDs

```

```{r analysis 1, include=TRUE}

# CODE PROVIDED BY AUTHORS IN TECHNICAL APPENDIX

# Before undertaking any descriptive statistics of the metric dimensions of the lances, it is important to re-sort the data matrix and subset the data.
lances_M$site<-factor(lances_M$site,
                    levels = c("Ejsbol","Nydam", "Illerup"),
                    labels = c("Ejsbol","Nydam", "Illerup"),
                    ordered=T)                                # This simply converts the variable 'site' into an ordered variable of factors with levels.

# The data can be subset into specific features of the lance, so that only cases with complete data are used
blade<-subset(lances_M, point=='yes')                                 # Selects all cases with complete blades
socket<-subset(lances_M, socket=='yes')                               # Selects all cases with complete sockets
total<-subset(lances_M, point=='yes' & socket=='yes')                 # Selects all 'complete' lances with all features

# Pairwise correlations of lance features
cor(socket$socket_length,socket$sock_bdiam,use="pairwise.complete.obs") # 0.52
cor(socket$sock_bdiam,socket$sock_sdiam,use="pairwise.complete.obs")    # 0.52
cor(lances_M$width,lances_M$thickness,use="pairwise.complete.obs")      # 0.53
cor(blade$blade_length,blade$width,use="pairwise.complete.obs")         # 0.58
cor(blade$blade_length,blade$thickness,use="pairwise.complete.obs")     # 0.47

# GPA is performed on the lances, including the semi-landmarks at the socket.
lancespr<-gpagen(lances_geo, print.progress = FALSE)  # GPA of all lances

# Manually create a vector containing site
site<-as.factor(c(rep("Ejsbøl",6),rep("Nydam",24),rep("Illerup",48)))
site<-factor(site,levels=c("Ejsbøl","Nydam","Illerup")) # Create a vector for the different sites

# Create a geomorph data frame for analysis
lances_geomorph<-geomorph.data.frame(shape=lancespr$coords,site=site,size=lancespr$Csize,ind=lances_GMM$site_lance.id) # Parse the data into a geomorph data frame
```

With the R session properly configured, correct packages installed, and data loaded and shaped according to the paper and provided R file, we were then able to proceed with our attempt at replicating the correlation matrix of lance dimensions shown in Figure 8 of the original paper. This is where we encountered our first issue, as the code used to create the figure was not included in the R file. Instead a series of five corr() statements (shown below) comparing a subset of lance features were included with the resulting correlation coefficients included in the comments.

Pairwise correlations of lance features
cor(socket$socket_length,socket$sock_bdiam,use="pairwise.complete.obs") # 0.52
cor(socket$sock_bdiam,socket$sock_sdiam,use="pairwise.complete.obs")    # 0.52
cor(lances_M$width,lances_M$thickness,use="pairwise.complete.obs")      # 0.53
cor(blade$blade_length,blade$width,use="pairwise.complete.obs")         # 0.58
cor(blade$blade_length,blade$thickness,use="pairwise.complete.obs")     # 0.47

We then attempted to replicate the figure with the ggpairs() library and perform a visual comparison to Figure 8.

Finally, we attempted to replicate the original finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis. This test was described on page 44 of the original paper as well as in the R document and relies on the procD.allometry() function from the geomorph package.

## Results

Unfortunately, the pairwise correlation coefficients noted in the comments were not the same as the ones shown in Figure 8. This inconsistiency made it more difficult to confirm the success of our replication attempt.

Luckily, we were able to replicate the finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis on page 44 of the original paper

### Reproducing Figure 8 and associated analysis

The figure that we are attepting to replicate is Figure 8, captioned as a  "Correlation matrix of Havor lance dimensions showing Pearson's r(lowerpanel), accompanying bivariate plots (upperpanel) and  distributions (histograms with kernel density estimation, on the diagonal). Red=Ejsbøl, green=Nydam, blue=Illerup."

Data appear to be drawn from the main lances_M dataframe, but the figure labels include "socket (max)" and "socket (min)" which do not haave obvious analogues in the column names.

A closer review of the paper leads us to the following line: "Although socket 'diameters' were measured, they are referred to here as minimum and maximum socket 'thickness' (p.38." This points us to two fields, which we can test against the summary table that is Table 1:

```{r replication target preparation, include=TRUE, echo=FALSE}

lances_M %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean=mean(sock_sdiam, na.rm = TRUE), n = sum(!is.na(sock_sdiam)))

lances_M %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean=mean(sock_bdiam, na.rm = TRUE), n = sum(!is.na(sock_bdiam)))

blade %>%
  dplyr::group_by(site) %>%
  dplyr::summarise(mean=mean(blade_length, na.rm = TRUE), n = sum(!is.na(blade_length)))
```

We can be confident therefore that "sock_bdiam" and "sock_sdiam" are the sources of "socket (max) [thickness (mm)]" and "socket (min) [thickness (mm)]" respectively. Using some custom functions and the excellent ggally package, we can produce a first replication of the figure.

```{r replication figure, include=TRUE, echo=FALSE}

#notes for those who follow:
# check variable names - dont think sock_bdiam etc. is the same as min max width
# color match
# label match

#upper_fn <- function(data, mapping, ...){
#  p <- ggplot(data = data, mapping = mapping) + 
#    geom_smooth(method=loess, fill="None", color="Black", ...) +
#    geom_point(aes(color=site, alpha=0.5))
#  p
#}

#REMOVED fill="None"
upper_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_smooth(method=loess, color="Black", ...) +
    geom_point(aes(color=site, alpha=0.5))
  p
}

diag_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_histogram(bins=10, color="black", fill="cyan", aes(y = ..ncount..)) + 
    geom_density(color="red",aes(y = ..scaled..,color="red"))
  p
}

pairs_fig <- ggpairs(lances_M, columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() #+
        #ggsave(here("Images","replication_fig_8v1.png"))
print(pairs_fig, progress = F)

#Copied to try to adjust to get to knit in Thomas's env
#pairs_fig <- ggpairs(lances_M, columns = c('total_length', 'blade_length', 'socket_length', 'sock_bdiam', #'sock_sdiam', 'width', 'thickness'),  
#        upper=list(continuous = upper_fn),
#        diag=list(continuous = diag_fn),
#        lower=list(continuous = "cor")) + theme_minimal() #+
#        #ggsave(here("Images","replication_fig_8v1.png"))
#print(pairs_fig, progress = F)

```

The correlation values to do not match the paper. A quick test with the stats passage verifies the results of our plot and confirms that the issue is not a mis-labelling of a spearman coefficient.

```{r replication corr test, include=TRUE, echo=FALSE}
cor(lances_M$socket_length,lances_M$sock_bdiam, use= "pairwise.complete.obs", method = "pearson")
cor(lances_M$socket_length,lances_M$sock_bdiam, use= "pairwise.complete.obs", method = "spearman")

cor(lances_M$sock_sdiam,lances_M$width, use= "pairwise.complete.obs", method = "pearson")
cor(lances_M$sock_sdiam,lances_M$width, use= "pairwise.complete.obs", method = "spearman")
```
Try recreating figure on different data subsets:

```{r replication corr test 2, include=TRUE, echo=FALSE}
#COMMENTED OUT BY THOMAS IN ORDER TO KNIT

pairs_fig <- ggpairs(lances_M[complete.cases(lances_M),], columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() #+
#        ggsave(here("Images","replication_fig_8_complete_cases_manual.png"))
print(pairs_fig, progress = F)

#COMMENTED OUT BY THOMAS IN ORDER TO KNIT
pairs_fig <- ggpairs(blade, columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() #+
#        ggsave(here("Images","replication_fig_8_blade.png"))
print(pairs_fig, progress = F)

#COMMENTED OUT BY THOMAS IN ORDER TO KNIT
pairs_fig <- ggpairs(socket, columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() #+
#        ggsave(here("Images","replication_fig_8_socket.png"))
print(pairs_fig, progress = F)

#COMMENTED OUT BY THOMAS IN ORDER TO KNIT
pairs_fig <- ggpairs(total, columns = c("total_length", "blade_length", "socket_length", "sock_bdiam", "sock_sdiam", "width", "thickness"), 
        upper=list(continuous = upper_fn),
        diag=list(continuous = diag_fn),
        lower=list(continuous = "cor")) + theme_minimal() #+
#        ggsave(here("Images","replication_fig_8_blades_and_sockets.png"))
print(pairs_fig, progress = F)

```

None of the author-defined data subsets (or a manual calculation based on on complete cases)allows us to replicate the correaltion results.

### Reproducing ANOVA analysis

As of geomorph version 3.1, the function procD.allmoetry has been deprecated (see https://rdrr.io/cran/geomorph/man/procD.allometry.html). This function underpinned the ANOVA tests provided in the technical code appendix. Thankfully, the specificity of the authors' instructions allow a previous version of geomorph to be installed!

We replicate the results of p.44 on the ANOVA tests used to establish whether there is a significant relationship between centroid size and shape.

```{r replication ANOVA  test, include=TRUE, echo=FALSE}
lances_ANOVA_site<-procD.allometry(shape ~ size, ~site, print.progress = FALSE,  data=lances_geomorph,method="PredLine")

print(lances_ANOVA_site)

```

The results match. We note that no seed is specified in the provided code, making the random resampling approach used volatile.

We can explore the effect of other parameters in the allometry ANOVA, to test whether the core conclusion holds. In the example below, we do not log-transform the size variable and we set a signficance threshold (alpha) of 0.1, rather than the default 0.05.

```{r explore ANOVA  test v1, include=TRUE, echo=FALSE}
lances_ANOVA_two<-procD.allometry(shape ~ size, ~site, logsz=FALSE, seed=42, alpha=0.1, print.progress = FALSE,  data=lances_geomorph, method="PredLine")

print(lances_ANOVA_two)

```

The core result holds! We can push these boundaries futher; removing residual randomnizatio and using Cohen’s f-squared values as the random distribution to
estimate effect size.

```{r explore ANOVA  test v2, include=TRUE, echo=FALSE}
lances_ANOVA_three <- procD.allometry(shape ~ size, ~site, logsz=FALSE, seed=42, alpha=0.1, effect.type="cohen", RRPP=FALSE,  print.progress = FALSE, data=lances_geomorph, method="PredLine")

print(lances_ANOVA_three)

```

The core result still holds.

## Conclusion

Our attempt at a replication of two parts of the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres was an overall success despite some missing documentation and inconsistencies between the paper and provided code.

Specifically, the replication attempt of the figure showing the pair-wise correlation of all lance dimensions across the three underlying datasets (original study Fig 8) was used to confirm that the underlying data provided by the authors was the same as that used in the analysis. This replication attempt was complicated by the lack of the code used to generate the plot in the provided R file. Additionally, the correlation coefficients noted in the R file comments did not match the figure in the original paper. We were able to create a similar figure to the one in the original paper; but without a clear expected outcome, we were unable to confirm that the provided dataset was the same one used in the analysis.

The second part of the original paper we attempted to replicate was the original finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis. The code used to perform this analysis was included in the R file, and the expected results noted in the comments matched the original paper. This made it much easier to evaluate the accuracy of our replication. We ran the same analysis as noted in the provided R code and obtained the same results.

The combination of these two replication efforts combined to make this a successful replication attempt. Although our initial attempt cast doubt on the underlying data, the fact that the second replication matched the expected result suggests that the problem with the first attempt was a transcription error between code and paper and not an indication of bad underlying data.

## References

1. Li, X.J., Bevan, A.H., Martinón-Torres, M., Rehren, T., Cao, W., Xia, Y., Zhao, K., 2014.
Crossbows and imperial craft organisation: the bronze triggers of China's Terracotta
Army. Antiquity 88, 126–140.

2. Birch, T., & Martinon-Torres, M. (2018). Shape as a measure of weapon standardisation: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia. Journal of Archaeological Science, 101, 34-51. doi: https://doi.org/10.1016/j.jas.2018.11.002

### 2020-02-28 Data/yjasc_4911_Appendix_A.R

@@ -1,0 +1,287 @@
+#______________________________________________________________________
+# METRIC AND GEOMETRIC MORPHOMETRIC ANALYSIS OF HAVOR LANCES: R SCRIPT
+# By Thomas Edward Birch
+# Date: 04-10-2018
+#______________________________________________________________________
+
+                          #-#-#-#-#-#-#-#-#-#-#-
+                          #### INTRODUCTION ----
+                          #-#-#-#-#-#-#-#-#-#-#-
+
+# The purpose of this R script is to make it possible for readers to evaluate
+# the data manipulation procedures and statistical analyses that were performed,
+# and to easily reproduce the results published.
+#
+# The aim is to provide a clear and transparent script that is self-explanatory
+# through heavy annotation, and thus make it possible for readers to critically
+# evaluate the quality of analysis being conducted and reliability of results.
+#
+# The code provided in this script utilised freely available and open source
+# software (FOSS) packages and programmes. As such, the script remains free
+# and open for anyone to copy and/or update as necessary for their own projects.
+# Accreditation of the original source of any used/updated code would be kindly
+# appreciated, so that future readers may always be able to refer to the original
+# source for their satisfaction.
+
+
+                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+                          #### 1. LOADING LIBRARIES ----
+                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+
+# Load packages from library (assuming these are already installed...)
+library(geomorph) # geomorph 3.0.6
+library(shapes) # shapes 1.2.3
+
+                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+                          #### 2. READING IN THE DATA ----
+                          #-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-
+
+# Set working directory
+setwd("C:\\FILEPATH") # Make sure this is the directory where you have downloaded the lances_GMM.csv datafile
+
+
[…]

### 2020-03-09 analysis/paper.Rmd
Fixed typos

@@
 
-Next, we read in and shaped the data obtained from the ScieceDirect archive of the paper using the same code as the original paper, slightly modified to account for our modified approach to setting the working directory.
+Next, we read in and shaped the data obtained from the ScienceDirect archive of the paper using the same code as the original paper, slightly modified to account for our modified approach to setting the working directory.
 
@@
 
-Data appear to be drawn from the main lances_M dataframe, but the figure labels include "socket (max)" and "socket (min)" which do not haave obvious analogues in the column names.
+Data appear to be drawn from the main lances_M dataframe, but the figure labels include "socket (max)" and "socket (min)" which do not have obvious analogues in the column names.
 
@@
 
-The correlation values to do not match the paper. A quick test with the stats passage verifies the results of our plot and confirms that the issue is not a mis-labelling of a spearman coefficient.
+The correlation values do not exactly match that of the paper.
+A quick test with the stats passage verifies the results of our plot and confirms that the issue is not a mis-labelling of a spearman coefficient.
 
@@
 ```
-Try recreating figure on different data subsets:
+Below we recreate the figure on different data subsets:
 
@@
 
-None of the author-defined data subsets (or a manual calculation based on on complete cases)allows us to replicate the correaltion results.
+None of the author-defined data subsets (or a manual calculation based on on complete cases) allows us to replicate the correlation results.
 
@@
 This function underpinned the ANOVA tests provided in the technical code appendix.
-Thankfully, the specificity of the authors' instructions allow a previous version of geomorph to be installed!
+Thankfully, the specificity of the authors' instructions allowed a previous version of geomorph to be installed!
 
@@
 We can explore the effect of other parameters in the allometry ANOVA, to test whether the core conclusion holds.
-In the example below, we do not log-transform the size variable and we set a signficance threshold (alpha) of 0.1, rather than the default 0.05.
+In the example below, we do not log-transform the size variable and we set a significance threshold (alpha) of 0.1, rather than the default 0.05.
 
@@
 
-The core result holds! We can push these boundaries futher; removing residual randomnizatio and using Cohen’s f-squared values as the random distribution to
-estimate effect size.
+The core result holds! We can push these boundaries further; removing residual randomization and using Cohen’s f-squared values as the random distribution to estimate effect size.
 

### 2020-03-12 analysis/paper.Rmd
adding description and license

@@
 title: "Replication in Archaeological Science: Shape as a Measure of Weapon Standardisation"
-author: "Team Fresh"
+author: "Corey Christopherson, Zachary Garcia, Richard Todd, Thomas Winegarden"
 date: "2/24/2020"

### 2020-03-12 tests/testthat.R

@@ -1,0 +1,6 @@
+library(testthat)
+library(
+    Christopherson-Garcia-Todd-Winegarden-replication-project)
+
+test_check("
+    Christopherson-Garcia-Todd-Winegarden-replication-project")

### 2020-03-13 analysis/paper.Rmd
uncommenting install_version for shapes and geomorph
changing == to >=

@@
 library(devtools)
-#install_version("geomorph", version = "3.0.6", repos = "http://cran.rstudio.com")
+install_version("geomorph", version = "3.0.6", repos = "http://cran.rstudio.com")
 #install_version("geomorph", version = "3.0.6", repos = "http://cran.r-project.org")
@@
 
-#install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.rstudio.com")
+install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.rstudio.com")
 #install_version("shapes", version = "1.2.3", repos = "http://cran.rstudio.com/") #"http://cran.us.r-project.org")

### 2020-03-14 analysis/paper.Rmd
Added citation references to the original paper where the paper title and provided date are mentioned

@@
 
-In this work, we are attempting to confirm that the underlying data provided is the same as that used in the original study by replicating the figure showing the pair-wise correlation of all lance dimensions across the three underlying datasets (original study Fig 8).
+In this work, we are attempting to confirm that the underlying data provided [2] is the same as that used in the original study by replicating the figure showing the pair-wise correlation of all lance dimensions across the three underlying datasets (original study Fig 8).
 We are also attempting to replicate the original finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis.
@@
 
-Our attempt at a replication of two parts of the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres was an overall success despite some missing documentation and inconsistencies between the paper and provided code.
+Our attempt at a replication of two parts of the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres [2] was an overall success despite some missing documentation and inconsistencies between the paper and provided code.
 

### 2020-03-14 analysis/paper.Rmd
Cleaning up comments

@@
 ```{r set working directory, include=TRUE}
+#previous solutions used before moving to the here package commented out below
 #setwd(dirname(rstudioapi::getSourceEditorContext()$path))
 #getwd()
+
 library(here)
@@
 ```{r replication ANOVA  test, include=TRUE, echo=FALSE}
+#
 lances_ANOVA_site<-procD.allometry(shape ~ size, ~site, print.progress = FALSE,  data=lances_geomorph,method="PredLine")
@@
 ```{r explore ANOVA  test v1, include=TRUE, echo=FALSE}
+#
 lances_ANOVA_two<-procD.allometry(shape ~ size, ~site, logsz=FALSE, seed=42, alpha=0.1, print.progress = FALSE,  data=lances_geomorph, method="PredLine")
@@
 ```{r explore ANOVA  test v2, include=TRUE, echo=FALSE}
+#procD.allometry geomorph anova. needs geomorph 3.0.6!
 lances_ANOVA_three <- procD.allometry(shape ~ size, ~site, logsz=FALSE, seed=42, alpha=0.1, effect.type="cohen", RRPP=FALSE,  print.progress = FALSE, data=lances_geomorph, method="PredLine")

### 2020-03-15 analysis/paper.Rmd
move things around to simplify R CMD check

@@
 # Read the appendix datafile containing a matrix of dimensional data (metric) as "lances_M" 
-lances_M<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc2.csv"),header=TRUE)
+lances_M<-read.csv(here("analysis/Data","1-s2.0-S0305440318306757-mmc2.csv"),header=TRUE)
 
@@
 #lances_GMM<-read.csv("Data/1-s2.0-S0305440318306757-mmc3.csv",header=TRUE)
-lances_GMM<-read.csv(here("Data","1-s2.0-S0305440318306757-mmc3.csv"),header=TRUE)
+lances_GMM<-read.csv(here("analysis/Data","1-s2.0-S0305440318306757-mmc3.csv"),header=TRUE)
 

### 2020-03-15 tests/testthat.R

@@ -1,6 +1,0 @@
-library(testthat)
-library(
-    Christopherson-Garcia-Todd-Winegarden-replication-project)
-
-test_check("
-    Christopherson-Garcia-Todd-Winegarden-replication-project")

### 2020-03-15 analysis/paper.Rmd
final checklist

@@
 date: "2/24/2020"
-output: word_document
+output: bookdown::word_document2
 ---
@@
 
-However, in the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres, a new approach is proposed.
+However, in the paper _Shape as a measure of weapon standardization: From metric to geometric morphometric analysis of the Iron Age ‘Havor’ lance from Southern Scandinavia_ by Birch and Martinon-Torres [-@citationkey], a new approach is proposed.
 Instead of relying on dimensional comparison of different CVs, they propose using a Geomorphic Morphometric (GMM) analysis as a general measure of overall shape [2].
@@
 
-```{r replication corr test 2, include=TRUE, echo=FALSE}
+```{r replication-corr-test-2, include=TRUE, echo=FALSE, fig.cap="My figure"}
 #COMMENTED OUT BY THOMAS IN ORDER TO KNIT
@@
 ```
+
+- [ ] In Figure \@ref(fig:replication-corr-test-2) we can see... each plot to have it's own caption and don't forget cross-reference in the text.
+
+- [ ] Try the broom pkg to get tidy output from the ANOVAs so we can print pretty tables.
+
+- [ ] citations and reference list
 

### 2020-03-15 analysis/paper.Rmd
a few last edits

@@
 The original analysis was conducted in an unspecified version of R.
-However, specific versions of geomorph(v3.0.6) and shapes(v1.2.3), two of the most important packages, were directly stated.
+However, specific versions of geomorph (v3.0.6) and shapes (v1.2.3), two of the most important packages, were directly stated.
 Our reproduction attempt began with importing the packages explicitly identified in the paper as well as those required in the technical code.
@@
 
-Our next step was to set the working directory.
+<!-- Our next step was to set the working directory.
 Although a trivial step, the original code used setwd() and a static path string while our reproduction attempt opted to use the more robust here() library as shown below.
+-->
 
@@
 
-Next, we read in and shaped the data obtained from the ScienceDirect archive of the paper using the same code as the original paper, slightly modified to account for our modified approach to setting the working directory.
+<!-- Next, we read in and shaped the data obtained from the ScienceDirect archive of the paper using the same code as the original paper, slightly modified to account for our modified approach to setting the working directory.
+-->
 
@@
 
-With the R session properly configured, correct packages installed, and data loaded and shaped according to the paper and provided R file, we were then able to proceed with our attempt at replicating the correlation matrix of lance dimensions shown in Figure 8 of the original paper.
-This is where we encountered our first issue, as the code used to create the figure was not included in the R file.
-Instead a series of five corr() statements (shown below) comparing a subset of lance features were included with the resulting correlation coefficients included in the comments.
+<!--With the R session properly configured, correct packages installed, and data loaded and shaped according to the paper and provided R file, we were then able to proceed with --> We focused out our replication on correlation matrix of lance dimensions shown in Figure 8 of the original paper.
+The R code used to create the figure was not included in the R file provided by the authors with their publication.
+Instead, the authors provided a series of five `corr()` statements (shown below) comparing a subset of lance features were included with the resulting correlation coefficients included in the comments.
 
@@
 
-Unfortunately, the pairwise correlation coefficients noted in the comments were not the same as the ones shown in Figure 8.
-This inconsistiency made it more difficult to confirm the success of our replication attempt.
+The pairwise correlation coefficients noted in the comments in the code provided by the authors were not the same as the ones shown in their Figure 8 of their published paper.
+This inconsistency made it more difficult to confirm the success of our replication attempt.
 
-Luckily, we were able to replicate the finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis on page 44 of the original paper
+We were able to replicate the finding that there was found to be an association between overall centroid size and shape that could not be explained through site difference via an ANOVA analysis on page 44 of the original paper
 
@@
 
-The figure that we are attepting to replicate is Figure 8, captioned as a  "Correlation matrix of Havor lance dimensions showing Pearson's r(lowerpanel), accompanying bivariate plots (upperpanel) and  distributions (histograms with kernel density estimation, on the diagonal).
+The figure that we are attempting to replicate is Figure 8, captioned as a  "Correlation matrix of Havor lance dimensions showing Pearson's r(lowerpanel), accompanying bivariate plots (upperpanel) and  distributions (histograms with kernel density estimation, on the diagonal).
 Red=Ejsbøl, green=Nydam, blue=Illerup."
@@
 The correlation values do not exactly match that of the paper.
-A quick test with the stats passage verifies the results of our plot and confirms that the issue is not a mis-labelling of a spearman coefficient.
+A test with the stats package verifies the results of our plot and confirms that the issue is not a mis-labelling of a Spearman coefficient.
 
@@
 ```
+
 Below we recreate the figure on different data subsets:
@@
 This function underpinned the ANOVA tests provided in the technical code appendix.
-Thankfully, the specificity of the authors' instructions allowed a previous version of geomorph to be installed!
+The specificity of the authors' instructions allowed a previous version of geomorph to be installed for this replication.
 
@@
 print(lances_ANOVA_site)
-
 ```
